\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2019-06-25 draft.tex
polish 4.2-5

\begin{wrapfigure}[11]{r}{3.5cm}
	\vspace{-17mm}
	\begin{algorithm}[H]
		\;\;$m := \cur + 10$; \;
		\Loop{}{
			$t:=\cur$\;
			\uIf{ $t> m$}{
				$m := t$ \;
			}
		};
		$out(m)$\;\;
	\end{algorithm}
	\vspace{-0.8cm}
	\caption{The max$^+$ reducer with input before the main loop.}
	\vspace{-12mm}
	\label{fig:reducer_max1}
\end{wrapfigure}

@@

\begin{wrapfigure}[12]{r}{5.3cm}
	\vspace{-9mm}
	\begin{algorithm}[H]
		$s:=1;$\;
		\Loop{}{
			\uIf{$s=1$}{$m := \cur + 10; s:= 2$}
			\uElse{
				$t:=\cur$\;
				\uIf{ $t > m$}{
					$m := t$ \;
				}
			}
		};
		$out(m)$
	\end{algorithm}	
	\vspace{-2mm}
	\caption{The max$^{+\mathtt{fix}}$ reducer.}
	\vspace{-4mm}
	\label{fig:reducer_max2}
\end{wrapfigure}

@@

After the transformation described above, the reducer $s_1;\rloop\{s_2\};s_3$
never calls the $\cur$ function in the $s_1$ part before entering the loop. W.o.l.g, we assume that the reducer reads one input $\cur$ in one loop iteration.
For the case that multiple reads from input occur in the loop, we can break the multiple reads to several single reads over multiple loop iterations using auxiliary variables and branch statements. 
The command $s_2$
can be viewed as a function $F$ that reads the values of all variables and the
current input before executing $s_2$ and outputs the values of all variables
after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1}
\rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$ where $n$ is
the current input value of the reducer, $x_i$ and $x_i'$ are the values of the
variables before and after the execution of $s_2$, respectively, for
$i\in[1,k]$. The construction of $F$ from $s_2$ can be done in the standard way.

We reduce the reducer commutativity checking to checking validity of the
following formula for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$:
\begin{equation} F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2,
F(n_1,x_1,x_2,\ldots,x_k) ). \label{eq:commu} \end{equation} Intuitively, the
formula says that starting from the same initial valuations of variables and
with two different input orders, $[n_1;n_2]$ and $[n_2;n_1]$, the values of all
program variables are the same after we execute the loop body twice. The first
execution reads $n_1$ and then reads $n_2$. The other execution reads the two
inputs in the reverse order. Since any permutation of the input can be obtained
by a sequence of permutations of neighbouring inputs, the validity of
Formula~\ref{eq:commu} implies that the permutations will not change the final
variable valuation and hence the output in $s_3$.

@@

In our initial experiments, we realised that Formula~\ref{eq:commu} is a
condition that is too strong for the reducer to be commutative. To illustrate
the issue, we present a simple example. In the loop body, the input is first
stored in a variable $t$ and this is then assigned to $s$, i.e., $t:=\cur;s:=t$.
After the loop, the value of $s$ is output. In this case, the $F$ function
returns the updated values of both $s$ and $t$ after the execution of the loop
body. Observe that $F(c_1, F(c_2,s,t)) = (c_1+c_2, c_1)$ and $F(c_2, F(c_1,s,t))
= (c_1+c_2, c_2)$. It follows that Formula~\ref{eq:commu} is invalid. The second
component of the returned tuples, which causes the invalidity, corresponds to
the value of $t$ after executing the loop body twice. Their values are $c_1$ and
$c_2$, respectively, in $F(c_1, F(c_2,s,t))$ and $F(c_2, F(c_1,s,t))$. However,
in this case, the value of $t$ will not affect the output of the reducer.

@@

In order to properly evaluate the performance of J-ReCoVer, we used the search
engine \emph{searchcode.com} to collect Java programs containing the key strings
``public void reduce'' or ``protected void reduce''. Since there is an upper
bound on the number of results returned from the search engine, we added
different search filters in order to get more data. We tried all $12$
combinations of six filters on the code length $\{<50$, $ 50\sim 250$, $250\sim
450$, $450\sim 650$, $650\sim 850$, $850\sim 1050$, $1050\sim 1250\}$ and two
filters for data sources $\{\url{github.com}, \url{bitbuckect.com}\}$. In total,
we got $11,346$ Java programs. We excluded cases that were not Hadoop MapReduce
reducer programs (those do not import the Hadoop library, do not extend or
implement the reducer interface) and obtained $1,273$ examples. We further
removed those that could not be compiled, those that were identical to each
other, and those with non-numerical data types (e.g., strings). We obtained
$118$ reducers as the final benchmarks. Table~\ref{tab:benchmark} contains more
details of the considered reducer functions.

>>>>

\begin{wrapfigure}[11]{r}{3.5cm}
	\vspace{-18mm}
	\begin{algorithm}[H]
		\;\;$m := \cur + 10$; \;
		\Loop{}{
			$t:=\cur$\;
			\uIf{ $t> m$}{
				$m := t$ \;
			}
		};
		$out(m)$\;\;
	\end{algorithm}
	\vspace{-0.8cm}
	\caption{The max$^+$ reducer with input before the main loop.}
	\vspace{-11mm}
	\label{fig:reducer_max1}
\end{wrapfigure}

@@

\begin{wrapfigure}[12]{r}{5.3cm}
	\vspace{-10mm}
	\begin{algorithm}[H]
		$s:=1;$\;
		\Loop{}{
			\uIf{$s=1$}{$m := \cur + 10; s:= 2$}
			\uElse{
				$t:=\cur$\;
				\uIf{ $t > m$}{
					$m := t$ \;
				}
			}
		};
		$out(m)$
	\end{algorithm}	
	\vspace{-2mm}
	\caption{The max$^{+\mathtt{fix}}$ reducer.}
	\vspace{-4mm}
	\label{fig:reducer_max2}
\end{wrapfigure}

@@

After the transformation described above, the reducer $s_1;\rloop\{s_2\};s_3$
never calls the $\cur$ function in the $s_1$ part before entering the loop. W.o.l.g, we assume that the reducer reads exactly one input in one loop iteration.
For the case that multiple reads from input occur in the loop, we can break the multiple reads to several single reads over multiple loop iterations using auxiliary variables and branch statements.

The command $s_2$
can be viewed as a function $F$ that reads the values of all variables and the
current input before executing $s_2$ and outputs the values of all variables
after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1}
\rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$ where $n$ is
the current input value of the reducer, $x_i$ and $x_i'$ are the values of the
variables before and after the execution of $s_2$, respectively, for
$i\in[1,k]$. The construction of $F$ from $s_2$ can be done in the standard way.

We reduce the reducer commutativity verification to checking validity of the
following formula for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$:
\begin{equation} F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2,
F(n_1,x_1,x_2,\ldots,x_k) ). \label{eq:commu} \end{equation} Intuitively, the
formula says that starting from the same initial valuations of variables and
with two different input orders, $[n_1;n_2]$ and $[n_2;n_1]$, the values of all
program variables are the same after we execute the loop body twice. The first
execution reads $n_1$ and then reads $n_2$. The other execution reads the two
inputs in the reverse order. Since any permutation of the input can be obtained
by a sequence of permutations of neighbouring inputs, the validity of
Formula~\ref{eq:commu} implies that the permutations will not change the final
variable valuation and hence the output in $s_3$.

@@

In our initial experiments, we realised that Formula~\ref{eq:commu} is a
condition that is too strong for the reducer to be commutative. To illustrate
the issue, we present a simple example. In the loop body, the input is first
stored in a variable $t$ and this is then assigned to $s$, i.e., $t:=\cur;s:=t$.
After the loop, output the value of $s$. In this case, the $F$ function
returns the updated values of both $s$ and $t$ after the execution of the loop
body. Observe that $F(c_1, F(c_2,s,t)) = (c_1+c_2, c_1)$ and $F(c_2, F(c_1,s,t))
= (c_1+c_2, c_2)$. It follows that Formula~\ref{eq:commu} is invalid. The second
component of the returned tuples, which causes the invalidity, corresponds to
the value of $t$ after executing the loop body twice. Their values are $c_1$ and
$c_2$, respectively, in $F(c_1, F(c_2,s,t))$ and $F(c_2, F(c_1,s,t))$. However,
in this case, the value of $t$ will not affect the output of the reducer.

@@

In order to properly evaluate the performance of J-ReCoVer, we used the search
engine \emph{searchcode.com} to collect Java programs containing the key strings
``public void reduce'' or ``protected void reduce''. Since there is an upper
bound on the number of results returned from the search engine, we added
different search filters in order to get more data. We tried all $12$
combinations of six filters on the code length $\{<50$, $ 50\sim 250$, $250\sim
450$, $450\sim 650$, $650\sim 850$, $850\sim 1050$, $1050\sim 1250\}$ and two
filters for data sources $\{\url{github.com}, \url{bitbuckect.com}\}$. In total,
we got $11,346$ Java programs. We excluded cases that were not Hadoop MapReduce
reducer programs (those do not import the Hadoop library, do not extend or
implement the reducer interface) and obtained $1,273$ examples. We further
removed duplicates, those that could not be compiled, and those with non-numerical data types (e.g., strings). We obtained
$118$ reducers as the final benchmarks. Table~\ref{tab:benchmark} contains more
details of the considered reducer functions.
